\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Judson's Abstract Algebra: Chapter 6}
\date{}

\begin{document}
\maketitle

\section*{1}

Suppose that $G$ is a finite group with an element $g$ of order 5 and an element $h$ of order 7. Why must $|G| \geq 35$.

\vspace{\baselineskip}

Recall Corollary 6.6: Suppose that $G$ is a finite group and $g \in G$. Then the order of $g$ must divide the number of elements in $G$. Thus $|G|$ must be divisible by 5 and by 7. The smallest such positive integer that has this property is 35. This proves that $|G| \geq 35$.


\section*{2}

Suppose that $G$ is a finite group with 60 elements. What are the orders of possible subgroups of $G$?

\vspace{\baselineskip}

The possible orders of subgroups of $G$ are the divisors of $G$: 1,2,3,4,5,6,10,12,15,20,30,60.


\section*{4}

Prove or disprove: Every subgroup of the integers has finite order.

\vspace*{\baselineskip}

This statement is false. Consider the group $2\mathbb{Z} = \{ ..., -2, 0, 2, ... \}$. Note that $0 \in 2\mathbb{Z}$. Note that since the sum of two even numbers is even that $2\mathbb{Z}$ is closed. Notice that if $x \in \mathbb{Z}$ is even then $-x$ is even as well, this proves that all inverses are in $\mathbb{Z}$. This proves that $2\mathbb{Z}$ is a subgroup of $\mathbb{Z}$ with infinite order.


\section*{7}

Verify Euler's Theorem for $n=15$ and $a=4$.

\vspace{\baselineskip}

Recall Euler's Theorem: Let $a$ and $n$ be integers such that $n > 0$ and $\gcd(a,n) = 1$. Then $a^{\phi(n)} \equiv 1 \mod n$. Note that $\phi(15) = 8$. Note that $\gcd(15, 4) = 1$. Consider

$$4^8 = 65536 \equiv 1 \mod 15.$$


\section*{11}

Let $H$ be a subgroup of a group $G$ and suppose that $g_1, g_2 \in G$. Prove that the following conditions are equivalent.

$$g_1 H = g_2 H$$
$$H g_1^{-1} = H g_2^{-1}$$
$$g_1 H \subseteq g_2 H$$
$$g_2  \in g_1 H$$
$$g_1^{-1} g_2 \in H$$

\vspace{\baselineskip}

Assume $g_1 H = g_2 H$. Let $x \in H g_1^{-1}$. By assumption there exist $h_1 \in H$ such that $x = h_1 g_1^{-1}$. Consider that since

$$x^{-1} = g_1 h^{-1}$$

it is true that $x^{-1} \in g_1 H$. Therefore there exists $h_2^{-1} \in H$ such that $x^{-1} = g_2 h_2^{-1}$. It follows that $x = h_2 g_2^{-2}$. Hence $x \in H g_2^{-1}$. This shows $H g_1^{-1} \subseteq H g_2^{-1}$. The proof that $H g_2^{-1} \subseteq H g_1^{-1}$. Is identical.

Assume $Hg_1^{-1} = Hg_2^{-1}$. The proof is very similar to the last proof. Let $x \in g_1 H$. Then there exists $h_1 \in H$ such that $x = g_1 h_1$. Note that since $x^{-1} = h_1^{-1} g_{-1}^{-1}$ it is true that $x^{-1} \in H g_1^{-1}$. Therefore there exists $h_2^{-1} \in H$ such that $x^{-1} = h_2^{-1} g_2^{-1}$. Note that $x = g_2 h_2 \in g_2 H$. Therefore $g_1 H \subseteq g_2 H$.

Assume $g_1 H \subseteq g_2 H$. Note that for all $h_1 \in H$ there exists $h_2 \in H$. Such that $g_1 h_1 = g_2 h_2$. Note that $g_2 = g_1 h_1 h_2 ^{-1}$. Since $h_1 h_2^{-1} \in H$ it is true that $g_2 \in g_1 H$.

Assume $g_2  \in g_1 H$. There exists $h \in H$ such that $g_2 = g_1 h$. Notice $h = g_1^{-1} g_2 \in H$.

Assume $g_1^{-1} g_2 \in H$. Note that there exists $h_1 \in H$ such that $g_1^{-1} g_2 = h_1$ and hence $g_1 = g_2 h_1^{-1}$. Let $x \in g_1 H$. Then there exists $h_2 \in H$ such that $g_1 h_2 = x$. Then $x = g_2 h_1^{-1} h_2 \in g_2 H$. This proves $g_1 H \subseteq g_2 H$. The reverse proof is identical.


\section*{13}

What fails in the proof of Theorem 6.3 if $\phi : \mathcal{L}_H \rightarrow \mathcal{R}_H$ is defined by $\phi(gH) = Hg$.

\vspace{\baselineskip}

Lemma 6.1 does not apply to this $\phi$ in order to show that $\phi$ is well-defined or to show that $\phi$ is one-to-one.


\section*{14}

Suppose that $g^n = e$. Show that the order of $g$ divides $n$.

\vspace{\baselineskip}

Clearly $|g| \leq n$. If $|g| = n$, then clearly $|g|$ divides $n$. In the case where $|g| < n$ we will proceed by contradiction. Assume $|g|$ does not divide $n$. Applying the division algorithm yields $n = q|g| + r$ where $0 < r < |g|$. Consider

$$e = g^n = g^{q|g| + r} = g^{q|g|}g^r = eg^r = g^r$$

this is a a contradiction of $0 < r < |g|$, hence $|g|$ divides $n$.




\end{document}