\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Judson's Abstract Algebra: Chapter 9}
\date{}

\begin{document}
\maketitle

\section*{1}

Prove that $\mathbb{Z} \cong n \mathbb{Z}$ for $n \neq 0$.

\vspace{\baselineskip}

Let $n \in \mathbb{Z}, n \neq 0$. Consider the function $f : \mathbb{Z} \rightarrow n \mathbb{Z}$

$$f(x) = nx.$$

Let $x,y \in \mathbb{Z}$. Note that 

$$f(x+y) = n(x+y) = nx + ny = f(x) + f(y).$$

Assume $f(x) = f(y)$. Then

$$f(x) = f(y)$$
$$nx = ny$$
$$x = y$$

so $f$ is injective. Note that $f$ is surjective by definition. Since $f$ is a bijection and preserves the group operation $f$ is an isomorphism.


\section*{2} 

Prove that $\mathbb{C}^*$ is isomorphic to the subgroup of $GL_2(\mathbb{R}$ consisting of matrices of form

$$
  \begin{pmatrix}
    a & b \\
    -b & a
  \end{pmatrix}.$$
  
We will call the subgroup described above $M$. Consider a function $f : \mathbb{C}^* \rightarrow M$ defined by 

$$f(a+bi) = \begin{pmatrix}
    a & b \\
    -b & a
  \end{pmatrix}.$$
  
Let $a,b,c,d \in \mathbb{R}$ where $ab \neq 0$ and $cd \neq 0$. First note that the function preserves the group operation since

\begin{align*}
f((a+bi)(c+di)) &= f(ac-db + (db + da)i) \\
&= \begin{pmatrix}
    ac-db & cb + da \\
    cb + da & ac - db
  \end{pmatrix} \\
&= = \begin{pmatrix}
    a & b \\
    -b & a
  \end{pmatrix}
  = \begin{pmatrix}
    c & d \\
    -d & c
  \end{pmatrix} \\
&= f(a+bi)f(c+di)
\end{align*}

Note that $f$ is surjective because $a,b$ are arbitrary real numbers. Notice that $f$ is injective because of element-wise equality of matrices.


\section*{7}

Show that any cyclic group of order $n$ is isomorphic to $\mathbb{Z}_n$.

\vspace{\baselineskip}

Let $G$ be a cyclic group of order $n$ generated by $a$ and define a function $\phi : \mathbb{Z}_n \rightarrow G$ by $\phi(k) = a^k$ where $0 \leq k < n$. Let $x,y \in \mathbb{Z}_n$. 
Notice that since

$$\phi(x + y) = a^{x+y} = a^x a^y = \phi(x) \phi(y)$$

that $\phi$ preserves the group operation. Assume $\phi(x) = \phi(y)$. Then

$$\phi(x) = \phi(y)$$
$$a^x = a^y$$
$$a^x a^{-y} = e$$
$$a^{x-y} = e.$$

Hence $x - y \equiv 0 \mod n$ and therefore $x \equiv y \mod n$ and $\phi$ is injective. Let $g \in G$. Since $G$ is cyclic, there exists $k \in \mathbb{Z}, 0 \leq k < n$ such that

$$g = a^k = \phi(k)$$

therefore $\phi$ is surjective. This proves that $\phi$ is an isomorphism, hence any cyclic group of order $n$ is isomorphic to $\mathbb{Z}_n$.


\section*{9}

Let $G = \mathbb{R} \setminus \{ -1 \}$ and define a binary operation on $G$ by 

$$a * b = a + b + ab.$$

Prove that $G$ is a group under this operation. Show that $(G, *)$ is isomorphic to the multiplicative group of nonzero real numbers.

\vspace{\baselineskip}

Let $a,b,c \in G$. Note that $G$ contains an identity since

$$a * 0 = a + 0 + 0a = a.$$

Note $G$ contains inverses since

\begin{align*}
a * \frac{-a}{1+a} &= a + \frac{-a}{1+a} + \frac{-a^2}{1+a} \\
&= \frac{a + a^2}{1+a} + \frac{-a}{1+a} + \frac{-a^2}{1+a} \\ 
&= 0.
\end{align*}

Consider

\begin{align*}
(a*b) * c &= (a + b + ab) * c \\
&= a + b + ab + c + c (a + b + ab) \\
&= a + b + c + ab + ac + bc + abc \\
&= b + c + bc + a + a (b + c + bc) \\
&= a * (b + c + bc) \\
&= a * (b*c)
\end{align*}

shows that $*$ is associative. We must now find a bijection $f : G \rightarrow \mathbb{R} \setminus \{ 0 \}$. Let $f(x) = 1 + x$. Consider

\begin{align*}
f(a*b) &= f(a + b + ab) \\ 
&= (1 + a + b + ab) \\
&= (1 + a) (1 + b) \\
&= f(a) f(b).
\end{align*}

Hence $f$ preserves the group operation. It is clear that $f$ is injective since

\begin{align*}
f(a) &= f(b) \\
1 + a &= 1 + b \\
a &= b.
\end{align*}

It is clear that $f$ is surjective since for $x \in \mathbb{R} \setminus \{ 0 \}$

$$f(x - 1) = x$$

and $(x - 1) \in G$.



\section*{27}

Let $G \equiv H$. Show that if $G$ is cyclic, then so is $H$.

\vspace{\baselineskip}

Let $a$ be a generator of $G$. Then for all $g \in G$ there exists $k \in \mathbb{Z}$ such that $a^k = g$. Let $f: G \rightarrow H$ be an isomorphism. Let $h \in H$. There exists $g \in G$ such that

\begin{align*}
h &= f(g) \\
&= f(a^k) \\
&= f(a)^k.
\end{align*}

Notice that $f(a)$ is a generator of $H$ since $h$ is arbitrary. This proves that $H$ is a cyclic group.


\section*{28}

Prove that any group $G$ of order $p$, $p$ a prime, must be isomorphic to $\mathbb{Z}_p$.

\vspace{\baselineskip}

Recall Theorem 9.3: If $G$ is a cyclic group of order $n$, then $G$ is isomorphic to $\mathbb{Z_n}$. Also recall Corollary 6.7: Let $|G| = p$ with $p$ a prime number. Then $G$ is cyclic and any $g \in G$ such that $ \neq e$ is a generator. The desired result follows immediately from combining these two results.


\section*{34}

An \textit{automorphism} of a group $G$ is an isomorphism with itself. Prove that complex conjugation is an automorphism of the additive group of complex numbers; that is, show that the map $\phi(a+bi) = a - bi$ is an isomorphism from $\mathbb{C}$ to $\mathbb{C}$.

\vspace{\baselineskip}

Note that since

\begin{align*}
\phi((a+bi) + (c+di)) &= \phi((a+c) + (b+d)i) \\
&= (a+c) - (b+d)i \\
&= (a-bi) + (c-di) \\
&= \phi(a+bi) + \phi(c+di)
\end{align*}

we know that $\phi$ preserves the group operation. The proof that $\phi$ is surjective is trivial (to get any element apply $\phi$ to the element's inverse). To show that $\phi$ is injective consider

$$\phi(a+bi) = \phi(c+di)$$
$$\overline{\phi(a+bi)} = \overline{\phi(c+di)}$$
$$a + bi = c + di.$$


\section*{35}

Prove that $a + ib \rightarrow a- ib$ is an automorphism of $\mathbb{C}^*$.

\vspace{\baselineskip}

Call the map $f$. Note that by problem 34 we know that $f$ is a bijection. To show that $f$ preserves the group operation consider

\begin{align*}
f((a+bi) (c+di)) &= f((ac - bd) + (ad + bc)i) \\
&= (ac - bd) - (ad + bc)i \\
&= (a-bi)(c-di) \\
&= f(a+bi) f(c+di).
\end{align*}


\section*{37}

We will denote the set of all automorphisms of $G$ by Aut$(G)$. Prove that Aut$(G)$ is a subgroup of $S_G$, the group of permutations of $G$.

\vspace{\baselineskip}

Note that the identity automorphism is the function $x \rightarrow x$. Clearly any automorphism composed with another automorphism results in an automorphism. Since automorphisms are bijections clearly inverses are contained within Aut$(G)$. This proves Aut$(G)$ is a subgroup of $S_G$.


\section*{50}

Prove that $A \times B$ is abelian if and only if $A$ and $B$ are abelian.

\vspace{\baselineskip}

Assume $A$ and $B$ are abelian. Let $(a_1, b_1), (a_2, b_2) \in A \times B$. Note that since

\begin{align*}
(a_1, b_1) (a_2, b_2) &= (a_1 a_2, b_1 b_2) \\
&= (a_2 a_1, b_2 b_1) \\
&= (a_2, b_2) (a_1, b_1)
\end{align*}

it is true that $A \times B$ is abelian. Assume that $A \times B$ is abelian. Let $(a_1, b_1), (a_2, b_2) \in A \times B$. Consider

\begin{align*}
(a_1, b_1) (a_2, b_2) &= (a_2, b_2) (a_1, b_1) \\ 
(a_1 a_2, b_1 b_2) &= (a_2 a_1, b_2 b_1).
\end{align*}

This shows that $a_1 a_2 = a_2 a_1$ and that $b_1 b_2 = b_2 b_1$, therefore $A$ and $B$ are abelian.


\end{document}